Optimal (Multiway) Spatial Joins

### 2024-05-19 msj.tex
yufei

SJs have been extensively studied in the database system , which has led to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance on real-world applications. We refer the interested readers to \cite{apr+00,mp01,???} as entry points into the literature.

@@

From the theory perspective, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem is solvable in $O(n \log n + \out)$ time by planesweep \cite{bcko08}. However, the problem becomes significantly more challenging as soon as $k$ reaches 3, i.e., the {\em multiway} scenario. All the solutions developed (see \cite{mp01,???} and the references therein) before 2022 all suffer from a worst-case time complexity of $O(n^k)$, namely, no better than the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ... \times R_k$.

@@

Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the problem $k$-SJ directly, imply algorithms for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.  Specifically, in \cite{ty22}, Tao and Yi studied multiple variants of interval intersection joins under updates. Most relevant to our context is the variant where the input contains, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ... \times \I_k$ such that $\bigcap_{i=1}^k I_i \neq \emptyset$. Set $n = \sum_{i=1}^k |\I_i|$. Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the new (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $\Delta$ is the number of such $k$-tuples (i.e., change in the result size). Tao and Yi \cite{ty22} presented a data structure to achieve the purpose with the space consumption of $O(n \polylog n)$. Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem  in $O((n + \out) \cdot \polylog n)$ time.

>>>>

SJs have been extensively studied in the database system , leading to the development of numerous interesting methods that, although lacking strong theoretical guarantees, demonstrate excellent empirical performance in real-world applications. We refer interested readers to \cite{apr+00,mp01,???} as entry points into the literature.

@@

From the perspective of theory, SJs are best understood when $k = 2$, i.e., the {\em pairwise} scenario, where the problem can be solved in $O(n \log n + \out)$ time by resorting to planesweep \cite{bcko08}. However, the problem becomes significantly more challenging as soon as $k$ reaches 3, i.e., the {\em multiway} scenario. All the solutions developed (see \cite{mp98,mp01,pmt99,???} and the references therein) before 2022 suffer from a worst-case time complexity of $O(n^k)$, offering no improvement over the naive method that simply enumerates the entire cartesian product $R_1 \times R_2 \times ... \times R_k$.

@@

Year 2022 witnessed two independent works \cite{ty22,kcko22} that, although not tackling the $k$-SJ problem directly, imply $k$-SJ algorithms with $O((n + \out) \cdot \polylog n)$ running time. Specifically, in \cite{ty22}, Tao and Yi studied several variants of ``interval intersection joins'' under updates. Most relevant to our context is the variant where the input includes, for each $i \in [k]$, a set $\I_i$ of 1D intervals in $\real$, and the join result comprises all $k$-tuples $(I_1, I_2, ..., I_k) \in \I_1 \times \I_2 \times ... \times \I_k$ with $\bigcap_{i=1}^k I_i \neq \emptyset$. Given the insertion (resp., deletion) of an interval in one of the $k$ sets, the objective is to list all the newly-appearing (resp., disappearing) $k$-tuples in the join result in $O((1+\Delta) \cdot \polylog n)$ time, where $n = \sum_{i=1}^k |\I_i|$ and $\Delta$ is the number of such $k$-tuples. Tao and Yi \cite{ty22} presented a data structure of $O(n \polylog n)$ space achieving the purpose. Combining their structure with the planesweep technique, one can obtain an algorithm for solving the $k$-SJ problem in $O((n + \out) \cdot \polylog n)$ time.
